BLAST: Block Applications for Things

### 2022-04-06 demo-eswc22/paper.tex
style guide says FAU in german
final version of ESWC 22 paper
improve paper according to reviewers’ comments

\begin{document}
%
\title{BLAST: Block Applications for Things}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Michael Freund\inst{1}\orcidID{0000-0003-1601-9331} \and
Thomas Wehr\inst{1}\orcidID{0000-0002-0678-5019} \and
Andreas Harth\inst{1,2}\orcidID{0000-0002-0702-510X}}
%
\authorrunning{M. Freund et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Fraunhofer Institute for Integrated Circuits IIS, Nürnberg, Germany 
\email{firstname.lastname@iis.fraunhofer.de}\\
\and
Friedrich–Alexander University Erlangen-Nürnberg, Nürnberg, Germany}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
  We introduce a block-based visual programming language called BLAST for programs involving connected devices with a Web of Things abstraction.
  We developed an editor and an execution environment for BLAST programs that runs in a web browser.
  We demonstrate that BLAST can be used to create programs that interact with a variety of devices.
  In particular we show the use of connected devices in a geofencing scenario.
\keywords{Block-based Programming \and Web of Things.}
\end{abstract}
%
%
%
\setcounter{footnote}{0}
\section{Introduction}
Connected devices form the basis for a wide variety of applications in industry settings, such as condition monitoring, energy consumption analysis, simulation and optimisation~\cite{CIMINO2019103130}.
%Given the multitude of protocols for interacting with such devices, application development is difficult.
The Web of Things (WoT)\footnote{\url{https://www.w3.org/WoT/}} aims at providing a set of standardised technologies to help simplify creating applications involving connected devices.
At the same time, casual users should be able to quickly program applications involving connected devices.
The programming languages currently in use for accessing devices are primarily text-based and are difficult to access for casual users.
One possible approach to improve accessibility and ease the creation of computer programs is the visual programming paradigm.

Visual programming does not require the user to enter any text; instead, programs are based on graphical elements like blocks, graphs, tables or diagrams.
Programs are created by arranging predefined blocks with images or text via a drag-and-drop interface.
Blocks can be geometrically aligned with other compatible blocks, like in a jigsaw puzzle, to form complex programs.
Since blocks can only be arranged in the correct way, no syntax errors can occur~\cite{10.11453341221}~\cite{maloney2010scratch}~\cite{10.1145/1089733.1089734}.

%The created block programs can be translated into source code of any predefined programming language and then executed.
The block-based approach can be used not only for learning programming concepts, but also for more complex problems.
Block languages have been successfully used in constructing SPARQL queries~\cite{7369012} or programming industrial robots~\cite{8120406}~\cite{tomlein2017visual}, both areas which normally require advanced programming skills.

@@

The editor uses the Blockly library and provides custom control and command blocks that make up the BLAST language.
Based on the user's arrangement of these blocks, the editor generates JavaScript code that can then be executed in the browser.
The system builds on the Web Bluetooth API\footnote{\url{https://webbluetoothcg.github.io/web-bluetooth/}} and the WebHID API\footnote{\url{https://wicg.github.io/webhid/}} to interact with devices.
The system uses the Web of Things abstraction as interface; we have implemented the WoT interface for eight devices that use the BLE GAP and GATT profiles as well as the USB and Bluetooth HID interface.
The WoT interface contains the metadata and interaction affordances as well as the JavaScript code needed for communication.
In addition, the system includes access to the camera and sound output via browser APIs using the WoT interface.
Further, the system includes access to the Web Speech API\footnote{\url{https://wicg.github.io/speech-api/}}.
The functionality to issue HTTP requests is based on the Fetch API\footnote{\url{https://fetch.spec.whatwg.org/}}, Social Linked Data (Solid) authentication is based on the Inrupt client libraries\footnote{\url{https://docs.inrupt.com/developer-tools/javascript/client-libraries/}}, and SPARQL query support is based on µRDF.js\footnote{\url{https://github.com/vcharpenay/uRDF.js/}}.
Next to the presented small scenario around geofences, we have also developed BLAST programs to control and monitor an automated picking line to prepare orders in a more complex logistics scenario.

@@

We have presented BLAST, a block-based language that simplifies the interaction with IoT devices, digital twins or other web resources without the need for a deeper understanding of different programming languages or network protocols.
We are currently working on an execution environment that runs on Node.js as a way to execute BLAST programs independently of browsers, which impose restrictions on accessing devices to provide security and privacy protection.
%This will allow BLAST to use the full Bluetooth specification and not only features exposed by the Web Bluetooth API, which restricts the set of accessible GATT attributes in a blocklist\footnote{\url{https://webbluetoothcg.github.io/web-bluetooth/\#the-gatt-blocklist}} to provide security and privacy protection.
%To this end, we are developing an execution environment that can load BLAST programs, convert them to JavaScript, and execute them as a server.
%In the future, it would enrich BLAST if there was a way to consume WoT TDs and derive interaction possibilities, which would allow BLAST to interact with any WoT device.
%@@@Web Bluetooth has certain restrictions (security, reading RSSI), server should not have these restrictions.

>>>>

\begin{document}
%
\title{BLAST: Block Applications for Things}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Michael Freund\inst{1}\orcidID{0000-0003-1601-9331} \and
Thomas Wehr\inst{1}\orcidID{0000-0002-0678-5019} \and
Andreas Harth\inst{1,2}\orcidID{0000-0002-0702-510X}}
%
\authorrunning{M. Freund et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Fraunhofer Institute for Integrated Circuits IIS, Nürnberg, Germany 
\email{firstname.lastname@iis.fraunhofer.de}\\
\and
Friedrich-Alexander-Universität Erlangen-Nürnberg, Nürnberg, Germany}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
  We introduce a block-based visual programming language called BLAST for programs involving connected devices with a Web of Things abstraction.
  We developed an editor and an execution environment for BLAST programs that runs in a web browser.
  We demonstrate that BLAST can be used to create programs that interact with a variety of devices.
  In particular we show the use of connected devices in a geofencing scenario.
\keywords{Block-based Programming \and Web of Things \and Graphical Programming.}
\end{abstract}
%
%
%
\setcounter{footnote}{0}
\section{Introduction}
Connected devices form the basis for a wide variety of applications in industry settings, such as condition monitoring, energy consumption analysis, simulation and optimisation~\cite{CIMINO2019103130}.
%Given the multitude of protocols for interacting with such devices, application development is difficult.
The Web of Things (WoT)\footnote{\url{https://www.w3.org/WoT/}} aims at providing a set of standardised technologies to help simplify creating applications involving connected devices.
At the same time, casual users should be able to quickly program applications involving connected devices.
The programming languages currently in use for accessing devices are primarily text-based and are difficult to access for casual users.
One possible approach to improve accessibility and ease the creation of computer programs is the visual programming paradigm.

Visual programming does not require the user to enter any text; instead, programs are based on graphical elements like blocks, graphs, tables or diagrams.
Programs are created by arranging predefined blocks with images or text via a drag-and-drop interface.
Blocks can be geometrically aligned with other compatible blocks, like in a jigsaw puzzle, to form complex programs.
Since blocks can only be arranged in the correct way, no syntax errors can occur~\cite{10.1145/1089733.1089734}~\cite{maloney2010scratch}~\cite{10.11453341221}.

%The created block programs can be translated into source code of any predefined programming language and then executed.
The block-based approach can be used not only for learning programming concepts, but also for more complex problems.
Block languages have been successfully used in constructing SPARQL queries~\cite{7369012} or programming industrial robots~\cite{tomlein2017visual}~\cite{8120406}, both areas which normally require advanced programming skills.

@@

The editor uses the Blockly library and provides custom control and command blocks that make up the BLAST language. The custom blocks and the BLAST environment in general follow all the important design principles introduced by Fraser \cite{fraser}, such as a clear border style or the avoidance of transitive connections.
Based on the user's arrangement of these blocks, the editor generates JavaScript code that can then be executed in the browser.
The system builds on the Web Bluetooth API\footnote{\url{https://webbluetoothcg.github.io/web-bluetooth/}} and the WebHID API\footnote{\url{https://wicg.github.io/webhid/}} to interact with devices.
BLAST uses the Web of Things abstraction as interface; we have implemented the WoT interface for eight devices that use the BLE GAP and GATT profiles as well as the USB and Bluetooth HID interface.
The WoT interface contains the metadata and interaction affordances as well as the JavaScript code needed for communication.
In addition, the system includes access to the camera and sound output via browser APIs using the WoT interface.
Further, the system includes access to the Web Speech API\footnote{\url{https://wicg.github.io/speech-api/}}.
The functionality to issue HTTP requests is based on the Fetch API\footnote{\url{https://fetch.spec.whatwg.org/}}, Social Linked Data (Solid) authentication is based on the Inrupt client libraries\footnote{\url{https://docs.inrupt.com/developer-tools/javascript/client-libraries/}}, and SPARQL query support is based on µRDF.js\footnote{\url{https://github.com/vcharpenay/uRDF.js/}}.
Next to the presented small scenario around geofences, we have also developed BLAST programs to accomplish five use cases\footnote{TODO: URI to videos} with varying complexity. Each use case introduces a different challenge that requiers the use of programming pattern in combination with the built-in blocks for properties, actions, events, requests and queries.

\section{Usability Evaluation }
There is a large number of standardized questionnaires in scientific literature that allow test users to evaluate the usability of a software product. We use the System Usability Scale (SUS) questionnaire to evaluate the user friendliness of BLAST. According to Assila et. al. \cite{assila2016standardized}, SUS provides reliable results over different sample sizes and converges fast to the correct conclusion.

We conducted an evaluation with nine participants. None of the participants had background knowledge regarding the Web of Things, RDF, or SPARQL. Participants were first asked to solve four tasks using BLAST, these included a simple "Hello World!" program, a periodic polling of RSSI data by using a loop, reading data from a connected IoT device, and as a final task, reading data and processing it using logical operators. After the successful completion of these tasks, the participants filled out the SUS questionnaires. The achieved SUS score across all questionnaires is 81.4 out of 100 possible points, which corresponds to good usability.

It should be noted that the survey was conducted with an older version of BLAST. However, the user interface has changed only minimally since then, which is why the results are still relevant.

@@

We have introduced BLAST, a block-based language that simplifies the interaction with IoT devices, digital twins or other web resources without the need for a deeper understanding of different programming languages or network protocols.
We are currently working on an execution environment that runs on Node.js as a way to execute BLAST programs independently of browsers, which impose restrictions on accessing devices to provide security and privacy protection.
%This will allow BLAST to use the full Bluetooth specification and not only features exposed by the Web Bluetooth API, which restricts the set of accessible GATT attributes in a blocklist\footnote{\url{https://webbluetoothcg.github.io/web-bluetooth/\#the-gatt-blocklist}} to provide security and privacy protection.
%To this end, we are developing an execution environment that can load BLAST programs, convert them to JavaScript, and execute them as a server.
%In the future, it would enrich BLAST if there was a way to consume WoT TDs and derive interaction possibilities, which would allow BLAST to interact with any WoT device.
%@@@Web Bluetooth has certain restrictions (security, reading RSSI), server should not have these restrictions.
